---
title: "Urban Disorder and the Seasons"
subtitle: "A Case Study of Five Countries"
author: "Allen Maggard"
institute: "SIS-750"
date: "2023/04/05 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: ["metropolis"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    
---
```{r setup, include=FALSE}
library(tidyverse)
library(scales)
library(haven)
library(dplyr)
library(patchwork)
library(lubridate)
library(knitr)

knitr::opts_chunk$set(
  echo = FALSE, cache = FALSE, prompt = FALSE,
  tidy = FALSE, comment = NA,
  message = FALSE, warning = FALSE,fig.width=11, fig.asp=0.590, fig.align='center', out.width='100%'
)

indevents2<-read_csv("indevents2.csv")
```

# About the Data
## Basics
- Urban Social Disorder v.3 (Peace Research Institute Oslo): 13,640 observations of urban unrest around the world between 1960 and 2014

- Focus on two variables: NPART (# of persons responsible for disorder) and NDEATH (# of persons who died as result of disorder)

- Research question: is urban disorder more likely to arise during certain months than during others?


## Particulars

- Data filtered to include instances of unrest for Bangladesh, India, Myanmar, Pakistan, and Sri Lanka

- Data further refined to exclude N/As and ambiguously coded values

- These refinements yielded 190 observations spanning 1960 and 2014

---

# Analysis 1: Number of Participants
- A note regarding NPART coding: 1 = less than 10; 2 = 10-100; 3 = 101-1,000; 
4 = 1,001-10,000; 5 = 10,001-100,000; 6 = 100,001-1,000,000; 7 = > 1,000,000
``` {r, include = FALSE}
indevents_alpha<-indevents2 %>% ggplot(aes(x = as.integer(BMONTH), y = NPART)) +
  geom_jitter(aes(color = COUNTRY)) + 
  geom_smooth(se = FALSE, color = "black") +
  scale_x_continuous(n.breaks = 12) +
  scale_y_continuous(n.breaks = 7) +
  labs(x = "Month in which disorder event took place", y = "Number of persons involved in disorder", color = "Location of disorder event")
```

``` {r, echo = FALSE}
plot(indevents_alpha)
```

---
# Analysis 2: Number of Deaths
``` {r, echo = FALSE}
indevents_beta <- indevents2 %>% ggplot(aes(x = as.integer(BMONTH), y = as.integer(NDEATH))) +
  geom_jitter(aes(color = COUNTRY)) + 
  geom_smooth(se = FALSE, color = "black") +
  scale_x_continuous(n.breaks = 12) +
  scale_y_continuous(n.breaks = 7) +
  labs(x = "Month in which disorder event took place", y = "Number of deaths during disorder event", color = "Location of disorder event") + 
  theme(legend.position = "none")


indevents_beta
```

